---
title: "Building Cohorts with STROBE Diagrams"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Building Cohorts with STROBE Diagrams}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 10,
  fig.height = 8
)
```

## Introduction

The STROBE (Strengthening the Reporting of Observational Studies in Epidemiology) statement provides guidelines for reporting observational studies. A key component is the flow diagram that shows how the study population was selected. The `strobe` package provides tools to create STROBE-compliant flow diagrams while building your cohort.

This vignette demonstrates how to use the functions in the **strobe** package to:

1. Initialize a cohort with inclusion criteria  
2. Apply sequential filters with exclusion tracking  
3. Review the complete filtering log

## Getting Started

```{r setup, message=FALSE, warning=FALSE, eval=requireNamespace("medicaldata", quietly = TRUE)}
library(strobe)
library(dplyr)
```

## Basic STROBE Workflow

The STROBE workflow follows these steps:

1. **Initialize** the cohort with `strobe_initialize()`  
2. **Filter** sequentially with `strobe_filter()`  
3. **Review** the log with `get_strobe_log()`

### Example: Cytomegalovirus

This example requires the `medicaldata` package from CRAN to demonstrate results.

```{r basic_example, eval=requireNamespace("medicaldata", quietly = TRUE)}

#Obtain data
library(medicaldata)
data(cytomegalovirus)

#Create cohort
cytomegalovirus %>%
  strobe_initialize(inclusion_label = "Initial transplant cohort") %>%
  strobe_filter(
    condition = "age >= 30",
    inclusion_label = "Age ≥ 30",
    exclusion_reason = "Excluded: Age < 30"
  ) %>%
  strobe_filter(
    condition = "recipient.cmv == 1",
    inclusion_label = "CMV positive recipients",
    exclusion_reason = "Excluded: CMV negative"
  ) %>%
  strobe_filter(
    condition = "prior.transplant == 0",
    inclusion_label = "No prior transplant",
    exclusion_reason = "Excluded: Prior transplant"
  )%>%
  dplyr::select(age, recipient.cmv, donor.cmv, prior.transplant, sex, race)%>%
  #Show first rows and selected columns of final DF
  head()
```

### Reviewing the Filtering Log

```{r review_log, eval=requireNamespace("medicaldata", quietly = TRUE)}
get_strobe_log()
```

The log shows:

- Each filtering step
- Number of records included/excluded at each step
- Cumulative counts
- Exclusion reasons

## Advanced Examples

### Complex Multi-Step Filtering

```{r complex_example, eval=requireNamespace("medicaldata", quietly = TRUE)}
cytomegalovirus %>%
  strobe_initialize(inclusion_label = "All transplant recipients") %>%
  strobe_filter(
    condition = "age >= 18 & age <= 65",
    inclusion_label = "Age 18–65",
    exclusion_reason = "Excluded: Outside 18–65"
  ) %>%
  strobe_filter(
    condition = "recipient.cmv == 1",
    inclusion_label = "CMV positive",
    exclusion_reason = "Excluded: CMV negative"
  ) %>%
  strobe_filter(
    condition = "donor.cmv == 1",
    inclusion_label = "Donor CMV positive",
    exclusion_reason = "Excluded: Donor CMV negative"
  ) %>%
  strobe_filter(
    condition = "prior.transplant == 0",
    inclusion_label = "No prior transplant",
    exclusion_reason = "Excluded: Prior transplant"
  )%>%
  #Show first rows and selected columns of final DF
  dplyr::select(age, recipient.cmv, donor.cmv, prior.transplant, sex, race)%>%
  head()
```

The filter log from the above pipeline is below:
```{r review_log_2, eval=requireNamespace("medicaldata", quietly = TRUE)}
get_strobe_log()
```

## See Also

- `?strobe_initialize`
- `?strobe_filter`
- `?get_strobe_log`
